Nail: A practical interface generator for data formats

### 2014-04-30 paper/osdi14/eval.tex
Design
ZIP grammar from hell

\section{Evaluation}
\label{s:eval}
\input{fig-ethernet}
\input{fig-dnsfull}
\input{fig-zip}
In our evaluation of Nail, we try to answer four questions:

@@

In particular, we show extracts from grammars for the following protocols: 
\begin{itemize}
\item DNS messages.
\item Ethernet packets containing ARP, ICMP and UDP.
\item ZIP files.
  See Figure~\ref{fig:zip-extract}.
\end{itemize}

\paragraph{DNS.}
Domain Name System(DNS) packets are described in RFC 1035 and our grammar almost directly
corresponds to the diagrams in Section 4 of that document. Each DNS packet consists of a header, a
set of question records and a set of answer records. Domain names in both queries and answers are
encoded as a sequence of Pascal-style strings. Each label is preceded by a length field, and the
domain name is terminated by a $0$ octet. The client's query packet contains one question
record to the server for each DNS look-up it wants to perform. In addition to the appropriate
response records, the server's reply contains a duplicate of the client's question section, which
allows stateless DNS clients and caches.

@@

\paragraph{Ethernet}
\XXX[TODO]
\paragraph{ZIP Files.}

Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}.
ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an \emph{end-of-directory
header}. This header contains a variable length comment, so it has to be located by scanning
backwards from the end of the file until a magic number and a valid length field is found. Many ZIP
implementations disagree on how to find this header in confusing situations, such as when the
comment contains the valid magic bytes\cite{wolf-berlinsides-zip}.
This end-of-directory header contains the offset and size of the \emph{ZIP directory}, which is an
array of one \emph{directory entry header} per file in the archive.
Each entry stores file meta data, such as file name, compressed and uncompressed size and a checksum,
in addition to the offset of a \emph{local file header}. The local file header duplicates most
information from the directory entry header and the compressed file contents follow it immediately.

@@

Furthermore, Nail transparently handles ZIP's data compression and checksums with stream
transformations. The implementation of one of these transforms are shown in
Figure~\ref{fig:zip-compress}. For formats such as Office documents or Android applications that are based on ZIP
containers, our ZIP grammar could be modified by including a more concrete grammar for file
contents, which currently are just an array of bytes.

With existing parser languages, the grammar can not guarantee correctness of checksums, length
fields or offsets, which places additional verification burden on the programmer. Furthermore,
existing parser generators cannot decompress data as they are parsing. This is merely inconvenient
for writing a grammar for plain ZIP files, but this prevents the grammar from validating the data
within a ZIP file.

The results suggest that Nail is a good fit for these data formats.

@@

\paragraph{Network stack.}
\XXX[Write this up]
\subsection{Security.} 
 \XXX[The ZIP format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous
security vulnerabilities in most implementations of ZIP and other archive formats. Furthermore,
Android security has been completely by-passed multiple times through parsing differentials in
handling ZIP files\cite{saurik-masterkey}. mention talk by Julia Wolf about zip]. 
Fuzz DNS. Fuzz ZIP. Design argument that security is hard to show.

>>>>

\section{Evaluation}
\label{s:eval}
\input{fig-ethernet}

In our evaluation of Nail, we try to answer four questions:

@@

In particular, we show extracts from grammars for the following protocols: 
\begin{itemize}
\item DNS messages.
\item Ethernet packets containing ARP, ICMP and UDP.
\item ZIP files.

\end{itemize}

\paragraph{DNS.}

\input{fig-dnsfull}
Domain Name System(DNS) packets are described in RFC 1035 and our grammar almost directly
corresponds to the diagrams in Section 4 of that document. Each DNS packet consists of a header, a
set of question records and a set of answer records. Domain names in both queries and answers are
encoded as a sequence of Pascal-style strings. Each label is preceded by a length field, and the
domain name is terminated by a $0$ octet. The client's query packet contains one question
record to the server for each DNS look-up it wants to perform. In addition to the appropriate
response records, the server's reply contains a duplicate of the client's question section, which
allows stateless DNS clients and caches.

@@

\paragraph{Ethernet}
\XXX[TODO]
\subsubsection{ZIP Files.}
\input{fig-zip}
\label{s:eval-format-zip}
Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}.
ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an \emph{end-of-directory
header}. This header contains a variable length comment, so it has to be located by scanning
backwards from the end of the file until a magic number and a valid length field is found. Many ZIP
implementations disagree on how to find this header in confusing situations, such as when the
comment contains the valid magic bytes\cite{wolf-berlinsides-zip}.
This end-of-directory header contains the offset and size of the \emph{ZIP directory}, which is an
array of one \emph{directory entry header} per file in the archive.
Each entry stores file meta data, such as file name, compressed and uncompressed size and a checksum,
in addition to the offset of a \emph{local file header}. The local file header duplicates most
information from the directory entry header and the compressed file contents follow it immediately.

@@

We have implemented a grammar for ZIP. A strongly abbreviated version is shown in
Figure~\ref{fig:zip-extract}.

Our ZIP grammar is a good example of transformations can capture complicated syntax in real-world
file format. The \texttt{zip\_file} grammar first splits the entire file stream into two streams
based on the \texttt{zip\_end\_of\_directory} transform.
\texttt{zip\_end\_of\_directory_parse(NailStream *files, NailStream *dir, NailStream (current))}
finds the end of directory header as described above by scanning the file backwards and splits the
file into two streams, one containing the end-of-directory header and one containing the file
contents. The header stream is then parsed. The file stream, to which all the offsets in the
metadata refer, is passed as an argument to the \texttt{end\_of\_directory} parser.
The directory header is then parsed and the offset and size transforms provided with Nail are used
to isolate the actual directory from the file stream. Each directory entry in turn points to a local
file header.

The \texttt{file} grammar for a zip file entry is particularly interesting. It receives the file
size, checksum and compression method of the file from the directory header. However, this same
information is duplicated in the file entry, so the grammar uses a standard \texttt{depend}
transform to check whether the two values are equal. Unlike most other transforms, \texttt{depend}
does not consume or produce strings, it only checks that two dependent fields are equal when parsing
and assigns the value of the second field to the first when generating. Immediately following the
file entry is the compressed data. Because most compression algorithms operate on unbounded streams
of data, Nail decompresses data in two steps.
First, it isolates the compressed data from the rest of the stream by using the size transform. In
this case, size operates on the current stream, which means it will consume data starting at the
current position of the parser in the input.
Second, Nail invokes a custom transform that implements the appropriate compression and decompression
functions based on the specified compression method. These functions are otherwise oblivious about
the layout or metadata of the file.

Existing parser languages are not able to handle a file format of this complexity.

\noindent The results suggest that Nail is a good fit for these data formats.

@@

\paragraph{Network stack.}
\XXX[Write this up]
\subsection{Security.} 
\XXX[Add PROTOS genome CVE's. count]
 \XXX[The ZIP format has been associated with many vulnerabilities, and the PROTOS Genome project found numerous
security vulnerabilities in most implementations of ZIP and other archive formats. Furthermore,
Android security has been completely by-passed multiple times through parsing differentials in
handling ZIP files\cite{saurik-masterkey}. mention talk by Julia Wolf about zip]. 
Fuzz DNS. Fuzz ZIP. Design argument that security is hard to show.

### 2014-04-30 paper/osdi14/intro.tex
ZIP grammar from hell

Third, many data formats contain redundancies, such as repeating information in multiple structures.
Applications usually do not explicitly check for consistency, and if different applications use
different instances of the same value, an attacker can craft input that causes inconsistencies.
Furthermore, security vulnerability can occur when an application assumes two repetitions of the
same data to be consistent.

>>>>

Third, many data formats contain redundancies, such as repeating information in multiple structures.
Applications usually do not explicitly check for consistency, and if different applications use
different instances of the same value, an attacker can craft input that causes inconsistencies.
Furthermore, security vulnerability can occur when an application assumes two repetitions of the
same data to be consistent, such as allocating a buffer based on one size field and copying into
that buffer based on another.
